Parsing Postfix log files

### 2009-03-05 doc/logparser-architecture.tex
Corrections to the first half of the architecture chapter.

    \item [Actions] Each action performs the work required to deal with a
        single \textit{category\/} of inputs, e.g.\ rejecting a delivery
        attempt.  Actions are invoked to process an input once it has been
        classified by a rule.

    \item [Rules]  The rules are responsible for classifying inputs; each
        rule recognises one input \textit{variant}, and specifies the
        action that will be invoked to process the input.  Each input
        category may have many input variants.  Rules provide an easily
        extensible method of associating inputs with actions.

@@

For each input the framework tries each rule in turn until it finds a rule
that recognises the input, then invokes the action specified by that rule.
If the input is not recognised by any of the rules, the framework issues a
warning; the framework can continue parsing, although for some inputs
stopping immediately may be preferable.

@@

Decoupling the parsing rules from their associated actions allows new rules
to be written and tested without requiring modifications to the parser
source code, significantly lowering the barrier to entry for casual users
who need to parse new inputs, e.g.\ part-time systems administrators
attempting to combat and reduce spam; it also allows companies to develop
user-extensible parsers without divulging their source code.  Decoupling
the framework, actions, and rules simplifies all three and creates a clear
separation of functionality: the framework manages the parsing process and
provides services to the actions, but does not need to perform anything
specific to the input being parsed; actions benefit from having services
provided by the framework, freeing them to concentrate on the task of
accurately and correctly processing the inputs and the information provided
by rules; rules are responsible for classifying inputs, and extracting data
from those inputs for processing by actions.

Separating the rules from the actions and framework makes it possible to
parse new inputs without modifying the core parsing algorithm.  Adding a
new rule with the action to invoke and a regex to recognise the inputs is
trivial in comparison to understanding an entire parser, identifying the
correct location to change, and making the appropriate changes.  Bear in
mind that changes to a parser must be made without adversely affecting
existing parsing, particularly as there may be edge cases that are not
immediately obvious --- \sectionref{yet-more-aborted-delivery-attempts}
describes a complication that occurs only four times in \numberOFlogFILES{}
log files.  Requiring changes to a parser's source code also complicates
upgrades of the parser, because the changes must be preserved during the
upgrade, and they may clash with changes made by the developer.  This
architecture allows the user to add new rules without having to edit a
parser, unless the new inputs cannot be processed by the existing actions.
If the new inputs do require new functionality, new actions can be added to
the parser without having to modify existing actions; only in the rare
event that the new actions need to cooperate with existing actions will
more extensive changes be required.

@@

Unlike traditional parsers such as those used when compiling a programming
language, this architecture does not require a fixed grammar specification
that inputs must adhere to.  The architecture is capable of dealing with
interleaved inputs, out of order inputs, and ambiguous inputs where
heuristics must be applied --- all have arisen and been successfully
accommodated in the Postfix log file parser.  This architecture is ideally
suited to parsing inputs where the input is not fully understood or does
not conform to a fixed grammar: the architecture warns about unparsed
inputs and other errors, but continues parsing as best it can, allowing the
developer of a new parser to decide which deficiencies are most important
and the order to address them in, rather than being forced to fix the first
error that arises.

@@

Most parsers will require the same basic functionality from the framework;
it is responsible for managing the parsing process from start to finish,
which generally will involve the following steps:

@@

    \item [Register actions]  Each action needs to be registered with the
        framework so that the framework knows about it; the list of
        registered actions will alsobe used when validating rules.

@@

            \item Ensuring the specified action has been registered.

            \item Checking for conflict in the data to be extracted, e.g.\
                setting the same variable twice.

            \item Checking that the regex used in the rule is valid.

@@

    \item [Classify the input]  The pseudo-code above shows how rules are
        successively tried until one is found that recognises the current
        input.  That pseudo-code is very simple: there may be efficiency
        concerns, (\sectionref{parser efficiency}), rule conditions
        (\sectionref{attaching conditions to rules}), or prioritisations
        (\sectionref{rules in architecture}) that complicate the process.

@@

    \item [Shared storage]  Parsers generally need to save some state
        information about the input being parsed, e.g.\ a compiler tracking
        which variables are in lexical scope as it moves from one lexical
        block to another.  The framework provides shared storage to deal
        with this and any other storage needs the actions have.  Actions
        may need to exchange data to correctly parse the input, e.g.\
        setting or clearing flags, maintaining a list of previously used
        identifiers, or ensuring at a higher level that the input being
        parsed meets a list of requirements.

    \item [Save and load state]  The architecture can save current state,
        viz the shared storage it provides for actions, and reload it
        later, so that information is not lost between parsing runs.

    \item [Specialised support functions]  Actions may need support or
        utility functions, and in general those functions should be
        available to all action.  Depending on the implementation the
        framework may be a good location for support functions; if there is
        another way to make those functions available to all actions it may
        be preferable to use it, maintaining a clear separation of
        concerns.

@@

Each action is a separate procedure written to process a particular
category of input, e.g.\ rejecting a delivery attempt.  There may be many
input variants within one input category; in general each action will
handle one category of inputs, with each rule recognising one input
variant.  It is anticipated that parsers based on this architecture will
have a high ratio of rules to actions, with the aim of having simple rules
and tightly focused actions.  An action may need to process different input
variants in slightly different ways, but large changes in processing
indicate the need for a new action and a new category of input; if an
action becomes overly complicated it starts to turn into a monolithic
parser, with too much logic contained in a single procedure.

The ability to easily add special purpose actions to deal with difficulties
and new requirements that are discovered during parser development is one
of the strengths of this architecture.  Instead of writing a single
monolithic function that processes every input and must be modified to
support new behaviour, with all the attendant risks of adversely affecting
the existing parsing, when a new requirement arises an independent action
can be written to satisfy it.  Sometimes the new action will require the
cooperation of other actions, e.g.\ to set or check a flag, so actions are
not always self-contained, but there will still be a far lower degree of
coupling and interdependency than in a monolithic parser.

During development of the Postfix log parser it became apparent that in
addition to the obvious variety in log~lines there were many complications
to be overcome.  Some were the result of deficiencies in Postfix's logging,
and some of those were rectified by later versions of Postfix, e.g.\
identifying bounce notifications
(\sectionref{identifying-bounce-notifications}); others were due to the
vagaries of process scheduling, client behaviour, and administrative
actions.  All were successfully accommodated in the Postfix log parser:
adding new actions was enough to overcome several of the complications;
others required modifications to a single existing action to work around a
difficulty; the remainder were resolved by adapting existing actions to
cooperate and exchange extra data (via the framework), changing their
behaviour as appropriate based on that extra data.  Every architecture
should aim to make the hard things possible as well as making the easy
things easy; the successful implementation of the Postfix log parser
demonstrates that this architecture achieves that aim.

@@

Actions may return a modified input line to the framework that will be
parsed as if read from the input stream, allowing for a simplified version
of cascaded parsing~\cite{cascaded-parsing}.  This powerful facility allows
several rules and actions to parse a single input, potentially simplifying
both rules and actions.

A simple example is to have one rule and action removing comments from
input lines, so that other rules and actions do not have to handle comments
at all; obviously if comment characters can be escaped or embedded in
quoted strings the implementation needs to be more complex, but for some
inputs this can greatly simplify subsequent parsing.

@@

A more complicated example is determining the atomic weight of a compound
given its chemical formula.  A chemical formula is a sequence of one or two
character chemical element abbreviations, each optionally followed by a
number indicating how many of that element are present in the formula (if
omitted the number is assumed to be one).  Some parts of a formula may be
repeated: they are enclosed in parentheses, and the closing parenthesis is
followed by a number indicating how many are present in the formula.  A
monomer is a simple compound, i.e.\ it does not have any repeated portions;
compounds with repeated portions are called polymers.  E.g.\ glucose (a
monomer) has the chemical formula $C_{6}H_{12}O_{6}$ and hydrocarbon (a
polymer) is $CH_{3}{(CH_{2})}_{50}CH_{3}$.  The atomic weight of an
individual element is given in periodic tables, but one element in
particular will be important in the later explanation: Hydrogen,
abbreviated H, has an atomic weight of 1.  The atomic weight of a given
compound is the sum of the atomic weights of each of the elements
multiplied by the number of times they occur; this is simple to calculate
for a monomer, but it is harder for a polymer because the atomic weight of
the repeated monomer must be multiplied by the number of times the monomer
is present in the compound (e.g.\ for hydrocarbon the atomic weight of
$CH_{2}$ must be multiplied by $50$).

Calculating the atomic weight of a monomer requires a single rule and
action: the action separates each element in the monomer, looks up its
atomic weight, multiplies it by the number of times it is present (one if
unspecified), and adds up the results.  The rule's regex is
\verb!^([A-Z]([a-z])?\d*)+$!, which matches:

    \item The start of the input.

    \item One or more of: an uppercase character, optionally followed by a
        lowercase character, followed by zero or more digits.

    \item The end of the input.

The regex does not guarantee that the chemical element abbreviations in the
formula are correct; the regex could be modified to check that by replacing
\verb![A-Z]([a-z])?! with all the valid abbreviations, slightly simplifying
the action at the cost of complicating the regex (which could be
automatically generated).

Parsing a polymer is more difficult because of the repeated monomers: each
monomer's atomic weight needs to be calculated, multiplied by the number of
times it is present in the compound, and that figure added to the sum.
Without cascaded parsing the action would need to maintain state while
recursively parsing the monomers embedded in the polymer.  With cascaded
parsing available three rules and actions could be used as shown in
\tableref{Regexes and Actions to calculate the atomic weight of a chemical
compound}:

@@

Rules are responsible for categorising inputs: each rule should recognise
one and only input variant.  An input category with multiple input variants
should have multiple rules, one for each variant.  Rules will typically use
a regex when classifying inputs, but other approaches may prove useful for
some applications, e.g.\ comparing fixed strings to the input, or checking
the length of the input; for the remainder of this thesis it will be
assumed that a regex is used.  Each rule must specify at least the regex to
recognise inputs and the action to invoke when classification is
successful, but implementations are free to add any other attributes they
require; \sectionref{rule attributes} describes the attributes used in
\parsername{}, and some potential attributes will be discussed later in
this section.

The architecture is line oriented at present: there is no facility for
rules to consume more input or push unused input back onto the input
stream.  This was not a deliberate design decision, rather a consequence of
the line oriented nature of Postfix log files; more flexible approaches
could be pursued.

Classifying inputs using the rules is simple: the first rule to recognise
the input determines the action that will be invoked; there is no
backtracking to try alternate rules; no attempt is made to pick a
\textit{best\/} rule.  \sectionref{attaching conditions to rules} contains
an example which requires that the rules are used in a specific order to
correctly parse the input, so a mechanism is needed to allow the author of
the rules to specify that ordering.  The framework should be written so
that it sorts the rules according to a priority attribute specified by the
rule author, allowing fine-grained control over the order that rules are
used in.  The priority attribute may be implemented as a number, or as a
range of values, e.g.\ low, medium, and high, or in a different fashion
entirely if it suits the implementation.  Rule ordering for efficiency is a
separate topic that is covered in \sectionref{rule ordering for
efficiency}; overlapping rules are discussed in \sectionref{overlapping
rules in architecture}.

@@

The framework issues a warning for each unparsed input, so it is evident
when the ruleset needs to be augmented.  Parsing new inputs is achieved in
one of three ways:

@@

    \item Write a new rule that pairs an existing action with a new regex;
        this adds a new variant to an existing category.

    \item Create a new category of inputs and a new action to process
        inputs from the new category; write a new rule pairing the new
        action with a new regex.

@@

Decoupling the rules from the actions and framework enables other rule
management approaches to be used, e.g.\ instead of manually editing
existing rules or adding new rules, machine learning techniques could be
used to automate the process.  If this approach was taken the choice of
machine learning technique would be constrained by the size of typical data
sets (see \sectionref{parser efficiency}).  Techniques requiring the full
data set when training would be impractical; Instance Based
Learning~\cite{instance-based-learning} techniques that automatically
determine which inputs from the training set are valuable and which inputs
can be discarded might reduce the data required to a manageable size.  A
parser could also dynamically create new rules in response to certain
inputs, e.g.\ parsing a subroutine declaration could cause a rule to be
created that parses calls to that subroutine, checking that the arguments
used agree with the subroutine's signature.

These avenues of research and development have not been pursued by the
author, but the architecture allows them to easily be undertaken
independently.

>>>>

    \item [Actions] Each action performs the work required to deal with a
        single \textit{category\/} of inputs, e.g.\ rejecting a delivery
        attempt.  Actions are invoked to process an input once it has been
        recognised by a rule.

    \item [Rules]  Rules are responsible for classifying inputs: each rule
        recognises one input \textit{variant\/} (a single input category
        may have many input variants).  Each rule also specifies the action
        to be invoked when an input has been recognised; rules thus provide
        an extensible method of associating inputs with actions.

@@

For each input the framework tries each rule in turn until it finds a rule
that recognises the input, then invokes the action specified by that rule.
If the input is not recognised by any of the rules, the framework issues a
warning; the framework can continue parsing after this, although for some
inputs stopping immediately may be preferable.

@@

Decoupling the parsing rules from their associated actions allows new rules
to be written and tested without requiring modifications to the parser
source code, significantly lowering the barrier to entry for casual users
who need to parse new inputs, e.g.\ part-time systems administrators
attempting to combat and reduce spam; it also allows companies to develop
user-extensible parsers without divulging their source code.  Decoupling
the framework, actions, and rules simplifies all three and creates a clear
separation of functionality: the framework manages the parsing process and
provides services to the actions; actions benefit from having services
provided by the framework, freeing them to concentrate on the task of
accurately and correctly processing the inputs and the information provided
by rules; rules are responsible for classifying inputs, and extracting data
from those inputs for processing by actions.

Separating the rules from the actions and framework makes it possible to
parse new inputs without modifying the core parsing algorithm.  Adding a
new rule with the action to invoke and a regex to recognise the inputs is
trivial in comparison to understanding an entire parser, identifying the
correct location to change, and making the appropriate changes.  Bear in
mind that changes to a parser must be made without adversely affecting
existing parsing, including any edge cases that are not immediately obvious
--- \sectionref{yet-more-aborted-delivery-attempts} describes a
complication that occurs only four times in \numberOFlogFILES{} log files.
Requiring changes to a parser's source code also complicates upgrades of
the parser, because the changes must be preserved during the upgrade, and
they may clash with changes made by the developer.  This architecture
allows the user to add new rules without having to edit a parser, unless
the new inputs cannot be processed by the existing actions.  If the new
inputs do require new functionality, new actions can be added to the parser
without having to modify existing actions; only in the rare event that the
new actions need to cooperate with existing actions will more extensive
changes be required.

@@

Unlike traditional parsers such as those used when compiling a programming
language, this architecture does not require a fixed grammar specification
that inputs must adhere to.  The architecture is capable of dealing with
interleaved inputs, out of order inputs, and ambiguous inputs where
heuristics must be applied --- all have arisen and been successfully
accommodated in \parsername{}.  This architecture is ideally suited to
parsing inputs where the input is not fully understood or does not conform
to a fixed grammar: the architecture warns about unparsed inputs and other
errors, but continues parsing as best it can, allowing the developer of a
new parser to decide which deficiencies are most important and the order to
address them in, rather than being forced to fix the first error that
arises.

@@

Most parsers will require the same basic functionality from the framework;
it is responsible for managing the parsing process from start to finish,
which will generally involve the following:

@@

    \item [Register actions]  Each action needs to be registered with the
        framework so that the framework knows about it; the list of
        registered actions will also be used when validating rules.

@@

            \item Ensuring the action specified by the rule has been
                registered with the framework.

            \item Checking for conflicts in the data to be extracted, e.g.\
                setting the same variable twice.

            \item Checking that the regex in the rule is valid.

@@

    \item [Classify the input]  The pseudo-code above shows how rules are
        successively tried until one is found that recognises the current
        input.  That pseudo-code is very simple: there may be efficiency
        concerns (\sectionref{parser efficiency}), rule conditions
        (\sectionref{attaching conditions to rules}), or prioritisations
        (\sectionref{rules in architecture}) that complicate the process.

@@

    \item [Shared storage]  Parsers commonly need to save some state
        information about the input being parsed, e.g.\ a compiler tracking
        which variables are in lexical scope as it moves from one lexical
        block to another.  The framework provides shared storage to deal
        with this and any other storage needs the actions have.  Actions
        may need to exchange data to correctly parse the input, e.g.\
        setting or clearing flags, maintaining a list of previously used
        identifiers, or ensuring at a higher level that the input being
        parsed meets requirements.

    \item [Save and load state]  The architecture can save its current
        state, i.e.\ the shared storage it provides for actions, and reload
        it later, so that information is not lost between parsing runs.

    \item [Specialised support functions]  Actions may need support or
        utility functions; the framework may be a good location for support
        functions, but if there is another way to make those functions
        available to all actions it may be preferable to use it instead,
        maintaining a clear separation of concerns.

@@

Each action is a separate procedure written to process a particular
category of input, e.g.\ rejecting a delivery attempt.  There may be many
input variants within one input category; in general each action will
handle one input category, with each rule recognising one input variant.
It is anticipated that parsers based on this architecture will have a high
ratio of rules to actions, with the aim of having simple rules and tightly
focused actions.  An action may need to process different input variants in
slightly different ways, but large changes in processing indicate the need
for a new action and a new category of input; if an action becomes overly
complicated it starts to turn into a monolithic parser, with too much logic
contained in a single procedure.

The ability to easily add special purpose actions to deal with difficulties
and new requirements that are discovered during parser development is one
of the strengths of this architecture.  Instead of writing a single
monolithic function that processes every input and must be modified to
support any new behaviour, with all the attendant risks of adversely
affecting the existing parsing, when a new requirement arises an
independent action can be written to satisfy it.  Sometimes the new action
will require the cooperation of other actions, e.g.\ to set or check a
flag, so actions are not always self-contained, but there will still be a
far lower degree of coupling and interdependency than in a monolithic
parser.

During development of \parsername{} it became apparent that in addition to
the obvious variety in log~lines there were many complications to be
overcome.  Some were the result of deficiencies in Postfix's logging, and
some of those deficiencies were rectified by later versions of Postfix,
e.g.\ identifying bounce notifications
(\sectionref{identifying-bounce-notifications}); others were due to the
vagaries of process scheduling, client behaviour, and administrative
actions.  All were successfully accommodated in \parsername{}: adding new
actions was enough to overcome several of the complications; others
required modifications to a single existing action to work around a
difficulty; the remainder were resolved by adapting existing actions to
cooperate and exchange extra data (via the framework), changing their
behaviour as appropriate based on that extra data.  Every architecture
should aim to make the easy things easy and the hard things possible; the
successful implementation of \parsername{} demonstrates that this
architecture achieves that aim.

@@

Actions may return a modified input line to the framework that will be
parsed as if read from the input stream, allowing for a simplified version
of cascaded parsing~\cite{cascaded-parsing}.  This powerful facility allows
several rules and actions to parse a single input, potentially simplifying
both rules and actions.  A simple example is to have one rule and action
removing comments from input lines, so that other rules and actions do not
have to handle comments at all; obviously if comment characters can be
escaped or embedded in quoted strings the implementation needs to be more
complex, but for some inputs this can greatly simplify subsequent parsing.

@@

A more complicated example is determining the atomic weight of a compound
given its chemical formula.  A chemical formula is a sequence of chemical
element abbreviations, each optionally followed by a number indicating how
many of that element are present in the formula (if omitted the number is
assumed to be one).  Some parts of a formula may be repeated: they are
enclosed in parentheses, and the closing parenthesis is followed by a
number indicating how many are present in the formula.  A monomer is a
simple compound, i.e.\ it does not have any repeated portions; compounds
with repeated portions are called polymers.  E.g.\ glucose, a monomer, has
the chemical formula $C_{6}H_{12}O_{6}$, and hydrocarbon, a polymer, is
$CH_{3}{(CH_{2})}_{50}CH_{3}$.  The atomic weight of an individual element
is given in periodic tables, but one element in particular will be
important in the later explanation: Hydrogen, abbreviated H, has an atomic
weight of 1.  The atomic weight of a given compound is the sum of the
atomic weights of each of its elements multiplied by the number of times
they occur; this is simple to calculate for a monomer, but it is harder for
a polymer because the atomic weight of the repeated monomer must be
multiplied by the number of times the monomer is present in the compound
(e.g.\ for hydrocarbon the atomic weight of $CH_{2}$ must be multiplied by
$50$).

Calculating the atomic weight of a monomer requires a single rule and
action.  The action separates each element in the monomer, looks up its
atomic weight, multiplies it by the number of times it is present (one if
unspecified), and adds up the results.  The rule's regex is
\verb!^([A-Z]([a-z])?\d*)+$!, which does not guarantee that the chemical
element abbreviations in the formula are correct; it could be modified to
check that by replacing \verb![A-Z]([a-z])?! with all the valid
abbreviations, slightly simplifying the action at the cost of complicating
the regex, which could be automatically generated.

Parsing a polymer is more difficult because of the repeated monomers: each
monomer's atomic weight needs to be calculated separately, multiplied by
the number of times it is present in the compound, and that figure added to
the sum.  Without cascaded parsing the action would need to maintain state
while recursively parsing the monomers embedded in the polymer.  Using
cascaded parsing, three rules and actions can be used instead as shown in
\tableref{Regexes and Actions to calculate the atomic weight of a chemical
compound}.

@@

Rules are responsible for categorising inputs: each rule should recognise
one and only one input variant; an input category with multiple input
variants should have multiple rules, one for each variant.  Rules will
typically use a regex when recognising inputs, but other approaches may
prove useful for some applications, e.g.\ comparing fixed strings to the
input, or checking the length of the input; for the remainder of this
thesis it will be assumed that a regex is used.  Each rule must specify at
least the regex to recognise inputs and the action to invoke when
recognition is successful, but implementations are free to add any other
attributes they require; \sectionref{rule attributes} describes the
attributes used in \parsername{}, and some generally useful attributes will
be discussed later in this section.

The architecture is line oriented at present: there is no facility for
rules to consume more input or push unused input back onto the input
stream.  This was not a deliberate design decision, rather a consequence of
the line oriented nature of Postfix log files; more flexible approaches
could be pursued.  Classifying inputs using the rules is simple: the first
rule to recognise the input determines the action that will be invoked;
there is no backtracking to try alternate rules; no attempt is made to pick
a \textit{best\/} rule.  \sectionref{attaching conditions to rules}
contains an example which requires that the rules are used in a specific
order to correctly parse the input, so a mechanism is needed to allow the
author of the rules to specify that ordering.  The framework should sort
the rules according to a priority attribute specified by the rule author,
allowing fine-grained control over the order that rules are used in.  The
priority attribute may be implemented as a number, or as a range of values,
e.g.\ low, medium, and high, or in a different fashion entirely if it suits
the implementation.  Rule ordering for efficiency is a separate topic that
is covered in \sectionref{rule ordering for efficiency}; overlapping rules
are discussed in \sectionref{overlapping rules in architecture}.

@@

The framework issues a warning for each unparsed input, so it is clearly
evident when the ruleset needs to be augmented.  Parsing new inputs is
achieved in one of three ways:

@@

    \item Write a new rule that pairs an existing action with a new regex,
        adding a new variant to an existing category.

    \item Create a new category of inputs and a new action to process
        inputs from the new category, and write a new rule pairing the new
        action with a new regex.

@@

Decoupling the rules from the actions and framework enables other rule
management approaches to be used, e.g.\ instead of manually editing
existing rules or adding new rules, machine learning techniques could be
used to automate the process.  If this approach was taken the choice of
machine learning technique would be constrained by the size of typical data
sets (see \sectionref{parser efficiency}).  Techniques requiring the full
data set when training would be impractical; Instance Based
Learning~\cite{instance-based-learning} techniques that automatically
determine which inputs from the training set are valuable and which inputs
can be discarded might reduce the data required to a manageable size.  A
parser could also dynamically create new rules in response to certain
inputs, e.g.\ parsing a subroutine declaration could cause a rule to be
created that parses calls to that subroutine, checking that the arguments
used agree with the subroutine's signature.  These avenues of research and
development have not been pursued by the author, but the architecture
allows them to easily be undertaken independently.

### 2009-03-05 doc/logparser-implementation.tex
Corrections to the first half of the architecture chapter.

    \item \acronym{SQL} is reasonably standard and many people will already
        be familiar with it; for those unfamiliar with it there are lots of
        readily available resources from which to learn (a good
        introduction to \acronym{SQL} can be found at
        \urlLastChecked{http://philip.greenspun.com/sql/}{2009/02/23}.
        Although every vendor implements a different dialect of
        \acronym{SQL}, the basics are the same everywhere (analogous to the
        overall similarities and minor differences amongst Irish English,
        British English, American English, and Australian English).
        Depending on the database in use there may be tools available that
        reduce or remove the requirement to know \acronym{SQL}; several are
        available for \gls{SQLite}, the default database used by the parser
        implementation.

>>>>

    \item \acronym{SQL} is reasonably standard and many people will already
        be familiar with it; for those unfamiliar with it there are lots of
        readily available resources from which to learn: a good
        introduction to \acronym{SQL} can be found at
        \urlLastChecked{http://philip.greenspun.com/sql/}{2009/02/23}.
        Although every vendor implements a different dialect of
        \acronym{SQL}, the basics are the same everywhere (analogous to the
        overall similarities and minor differences amongst Irish English,
        British English, American English, and Australian English).
        Depending on the database in use there may be tools available that
        reduce or remove the requirement to know \acronym{SQL}; several are
        available for \gls{SQLite}, the default database used by the parser
        implementation.
